---
title: "BankChurners"
author: "Jarod Wright"
date: "6/9/2021"
output: html_document
---
<style type="text/css">
  body{
  font-size: 18pt;
}
</style>

## Bank Churners

This is an analysis of churning customers. Each observation consists of 18 features and the set of all observations is very unbalanced. 


```{r , message=FALSE, warning=FALSE}

#dbplyr: for data stored in a relational database. Translates your dplyr code to SQL.(included in tidyverse)
# %>% pipe operator 

set.seed(123)

library(tidyverse)
library(caret)
library(caTools) #subset function
library(corrr)
library(RColorBrewer)
library(smotefamily)
library(plotly)
library(rcompanion)
library(corrplot)
library(randomForest)

coul1 <- brewer.pal(9, "Set3") 
coul2 <- brewer.pal(8, "Set2")


BankChurnersData <- read_csv("BankChurners.csv")
attach(BankChurnersData)
#Exclude last two columns and filter missing data
BankChurnersData <- BankChurnersData %>% select(c(1:21)) %>% arrange(Attrition_Flag)


```
Strength of association is calculated for nominal vs nominal with a bias corrected Cramer's V,  numeric vs numeric with Spearman (default) or Pearson correlation, and nominal vs numeric with ANOVA.

```{r , message=FALSE, warning=FALSE}

BinaryAssingment <- function(x){
  if(x=="Existing Customer"){
    return(0)
    
  } else {
    return(1)
  }
  
}

# Calculate a pairwise association between all variables in a data-frame. In particular nominal vs nominal with Chi-square, numeric vs numeric with Pearson correlation, and nominal vs numeric with ANOVA.
# Adopted from https://stackoverflow.com/a/52557631/590437
mixed_assoc = function(df, cor_method="spearman", adjust_cramersv_bias=TRUE){
  df_comb = expand.grid(names(df), names(df),  stringsAsFactors = F) %>% set_names("X1", "X2")
  
  is_nominal = function(x) class(x) %in% c("factor", "character")
  # https://community.rstudio.com/t/why-is-purr-is-numeric-deprecated/3559
  # https://github.com/r-lib/rlang/issues/781
  is_numeric <- function(x) { is.integer(x) || is_double(x)}
  
  f = function(xName,yName) {
    x =  pull(df, xName)
    y =  pull(df, yName)
    
    result = if(is_nominal(x) && is_nominal(y)){
      # use bias corrected cramersV as described in https://rdrr.io/cran/rcompanion/man/cramerV.html
      cv = cramerV(as.character(x), as.character(y), bias.correct = adjust_cramersv_bias)
      data.frame(xName, yName, assoc=cv, type="cramersV")
      
    }else if(is_numeric(x) && is_numeric(y)){
      correlation = cor(x, y, method=cor_method, use="complete.obs")
      data.frame(xName, yName, assoc=correlation, type="correlation")
      
    }else if(is_numeric(x) && is_nominal(y)){
      # from https://stats.stackexchange.com/questions/119835/correlation-between-a-nominal-iv-and-a-continuous-dv-variable/124618#124618
      r_squared = summary(lm(x ~ y))$r.squared
      data.frame(xName, yName, assoc=sqrt(r_squared), type="anova")
      
    }else if(is_nominal(x) && is_numeric(y)){
      r_squared = summary(lm(y ~x))$r.squared
      data.frame(xName, yName, assoc=sqrt(r_squared), type="anova")
      
    }else {
      warning(paste("unmatched column type combination: ", class(x), class(y)))
    }
    
    # finally add complete obs number and ratio to table
    result %>% mutate(complete_obs_pairs=sum(!is.na(x) & !is.na(y)), complete_obs_ratio=complete_obs_pairs/length(x)) %>% rename(x=xName, y=yName)
  }
  
  # apply function to each variable combination
  map2_df(df_comb$X1, df_comb$X2, f)
}
#------------------------------------------------------------
```
The correlation of all dependent variables are selected below. The most correlated features are used to generate models.

```{r , message=FALSE, warning=FALSE}

df.cor <- mixed_assoc(BankChurnersData)
Predictor.Correlation <- df.cor %>% filter(assoc > 0.5 & assoc != 1) # significant correlation between all variables
Sig.Corr <- df.cor %>% filter(x == "Attrition_Flag" | y == "Attrition_Flag") %>% filter(assoc != 1) %>% arrange(x, desc(assoc))# correlation of interest 


print(Sig.Corr[1:6,])



#correlation network graph, 
BankChurnersData %>%
  select(Total_Trans_Ct, Total_Ct_Chng_Q4_Q1, Total_Revolving_Bal, Contacts_Count_12_mon, Attrition_Flag) %>%
  mixed_assoc() %>%
  select(x, y, assoc) %>%
  spread(y, assoc) %>%
  column_to_rownames("x") %>%
  as.matrix %>%
  as_cordf %>%
  network_plot(min_cor = 0, colours = c("yellow", "blue", "red")) 
Data <- BankChurnersData %>% select(Attrition_Flag,Total_Trans_Ct,Total_Revolving_Bal,Contacts_Count_12_mon,Avg_Utilization_Ratio,Total_Trans_Amt)
Data$Attrition_Flag <- sapply(Data$Attrition_Flag, BinaryAssingment)

intrain      <- createDataPartition(Data$Attrition_Flag, p = 0.8, list = F)
imbal_train  <- Data[intrain, ]
imbal_test   <- Data[-intrain, ]

Churners <- BankChurnersData  %>% filter(Attrition_Flag == "Attrited Customer") 
Customers <- BankChurnersData %>% filter(Attrition_Flag == "Existing Customer")

```

# Plots 
```{r , message=FALSE, warning=FALSE}
Attritioncount <- table(BankChurnersData$Attrition_Flag) #unbalanced data
barplot(Attritioncount, col=coul1, main = "Proportion of Attrited to Exsiting")





Attrited.CC <- ggplot(BankChurnersData, aes(x=as.factor(Attrition_Flag), fill=as.factor(Card_Category))) + 
     geom_bar() + labs(x = " Attrition", fill = "Card Category") + scale_fill_brewer(palette = "Set2")
ggplotly(Attrited.CC) %>% config(displayModeBar = FALSE)

Attrited.Gender <- ggplot(BankChurnersData, aes(x=as.factor(Attrition_Flag), fill=as.factor(Gender))) + 
  geom_bar() + labs(x = " Attrition", fill = "Gender") + scale_fill_brewer(palette = "Set2")
ggplotly(Attrited.Gender) %>% config(displayModeBar = FALSE)

Attrited.Education <- ggplot(BankChurnersData, aes(x=as.factor(Attrition_Flag), fill=as.factor(Education_Level))) + 
  geom_bar() + labs(x = " Attrition", fill = "Education") + scale_fill_brewer(palette = "Set2")
ggplotly(Attrited.Education) %>% config(displayModeBar = FALSE)

Attrited.Marital <- ggplot(BankChurnersData, aes(x=as.factor(Attrition_Flag), fill=as.factor(Marital_Status))) + 
  geom_bar() + labs(x = " Attrition", fill = "Marital Status") + scale_fill_brewer(palette = "Set2")
ggplotly(Attrited.Marital) %>% config(displayModeBar = FALSE)



```
Synthetic minority oversampling is utilized to deal with the unbalance of the data.

```{r , message=FALSE, warning=FALSE}


#SMOTE 
smote_train <- SMOTE(imbal_train[,-1],imbal_train[,1] , K=5)
smote_train <- smote_train$data
smote_train$class <- as.factor(smote_train$class)



#logistic regression 
logit.model <- glm(class~.,data = smote_train, family = binomial)
logit.model_prob <- predict(logit.model, newdata=imbal_test,type = "response")
logit.pred <- as.factor(ifelse(logit.model_prob > 0.5, "Attrited Customer", "Existing Customer"))


#random forest
rf.model <- randomForest(class~., data= smote_train, ntree= 500, importance = TRUE)
rf.model_pred<- predict(rf.model, newdata = imbal_test, type = "response")
rf.model_pred<- factor(rf.model_pred, levels = c(0,1), labels=c("Existing Customer", "Attrited Customer"))


plot(rf.model)


imbal_test$Attrition_Flag <- factor(imbal_test$Attrition_Flag, 
                                    levels = c(0,1), labels = c("Existing Customer", "Attrited Customer"))

confusionMatrix(imbal_test$Attrition_Flag,logit.pred) ## poor performance
confusionMatrix(imbal_test$Attrition_Flag, rf.model_pred)


```


